---
title: "Model Selection"
author: "Jeetendra Gan | Student number - 50325023"
date: "10/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

Let's start by exploring the data a bit.

```{r}
library(ISLR)
summary(Hitters)
```

Here are the names of predictors and the response. We will try and predict the response, Salary of a player.

```{r}
names(Hitters)
```

There are some missing values here, so before we proceed we will remove them.
```{r}
Hitters = na.omit(Hitters)
```

Let us find the correlation matrix of the features.
```{r}
corrgram::corrgram(Hitters)
```

Let's see how we can do the best subset selection!
```{r}
library(leaps)
#regfit.full = regsubsets(Salary ~., data = Hitters)
#summary(regfit.full)
```

It only goes upto 9 variables by default, we want the best subset so let's look at each of the 19 variables.
```{r}
regfit.full = regsubsets(Salary~., data = Hitters, nvmax = 19)
reg.summary = summary(regfit.full)
reg.summary
names(reg.summary)
```

After running through the subset selection, we get 19 models which contain the best models with feaures ranging from 1 to 19. Here are plots for R-square, RSS, Adjusted R-Square values.
```{r}
features = 1:length(reg.summary$rsq)
plot(features, reg.summary$rsq, xlab = "Features", ylab="R^2",  type = "l")
points(10, reg.summary$rsq[10], pch=20, col="red")
plot(features, reg.summary$rss, xlab = "Features", ylab="RSS",  type = "l")
points(10, reg.summary$rss[10], pch=20, col="red")
plot(features, reg.summary$adjr2, xlab = "Features", ylab = "Adj. R^2",  type = "l")
fullSubStSummary = reg.summary;
```

We will select the model with 10 features due to the reasons mentioned below.
1. The RSS does not reduce much after 10.
2. The R^2 does not increase much after 10.
The above two observations indicate that there is not much gain by adding the 11th feature to the model.

```{r}
plot(regfit.full, scale="Cp")
coeffBestSubset = coef(regfit.full, 10)
```

### **Forward Stepwise selection**
____________________________________
```{r}
regfit.fwd = regsubsets(Salary ~ ., data = Hitters, nvmax = 19, method = "forward")
plot(regfit.fwd, scale="Cp")
fwdStSummary = summary(regfit.fwd)
```
Now, let us plot the R-Square, RSS, Adjusted R-Square for each of the feature subsets.
```{r}
x = 1:19
plot(x, fwdStSummary$rsq, xlab="Subset count",  ylab = "R-Square", type = "b", col="green")
plot(x, fwdStSummary$rss, xlab="Subset count", ylab = "RSS",  type = "b")
plot(x, fwdStSummary$adjr2, xlab = "Subset count", ylab = "Adj. R-Square", type = "b")
```

Now, let us plot BIC for both forward stepwise and best subset selection.
  
```{r}
rangeX = range(x)
rangeY = range(fullSubStSummary$bic)
plot(x, fullSubStSummary$bic, xlab = "Feature set", ylab="BIC", type="b", col="red", pch=24)
par(new = T)
rangeY = range(fwdStSummary$bic)
plot(x, fwdStSummary$bic, axes = FALSE, type="b", col="green", xlab = "", ylab="")
#ylim=range(rangeX, rangeY)
```
Now, let us look at the features selected in full subset selection and forward stepwise selection.

**For best subset selection**
```{r}
coeffFwdSubset = coef(regfit.full, 10)
coeffBestSubset
```
**For forward stepwise selection **
```{r}
coeffFwdSubset
```

In both the cases, same features are selected. They even have the same values.
